---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# cartomisc

<!-- badges: start -->
[![R build status](https://github.com/statnmap/cartomisc/workflows/R-CMD-check/badge.svg)](https://github.com/statnmap/cartomisc/actions)
<!-- badges: end -->

The goal of {cartomisc} is to store a few useful functions for spatial data manipulation and analysis. 

## Installation

You can install the dev version on Github

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("statnmap/cartomisc")
```

# Usage

```{r, message=FALSE}
library(dplyr)
library(raster)
library(cartomisc)
library(ggplot2)
library(sf)
```

## Extract part of the data with `gplot_data`

To draw multiple raster on the same ggplot2, or to draw raster after other objects (points, polygons, lines)

- data are extracted as a tibble
- data can then be added to a ggplot2 with `geom_tile`

```{r}
# Read some data
slogo <- stack(system.file("external/rlogo.grd", package = "raster")) 
slogo

# Get partial raster data to plot in ggplot
r.gg <- gplot_data(slogo)

# Remove NA to reduce size of table
r.gg.nona <- r.gg %>% 
  filter(!is.na(value))

r.gg.nona
```

### Plot with ggplot2

```{r, fig.width=12, fig.height=4}
# Plot
ggplot(r.gg.nona) +
  geom_tile(aes(x = x, y = y, fill = value)) +
  scale_fill_gradient("Probability", low = 'yellow', high = 'blue') +
  facet_wrap(vars(variable)) +
  coord_equal()
```

## Sun position

Calculate sun position for hillShade

```{r}
sun_position <- sun_position(2019, 04, 26,
  hour = 12, min = 0, sec = 0,
  lat = 46.5, long = 6.5
)
sun_position
```

Use with hillShade

```{r}
r <- raster(system.file("external/rlogo.grd", package = "raster"))
# slope and aspect
r_slope <- terrain(r, opt = "slope")
r_aspect <- terrain(r, opt = "aspect")
# hillshade
r_hillshade <- hillShade(r_slope, r_aspect, angle = sun_position$elevation, direction = sun_position$azimuth)
# plot
image(r)
image(r_hillshade, col = grey(seq(0, 1, 0.1), alpha = 0.5), add = TRUE)
```

## Create buffer areas with attribute of the closest region

- Download some data
```{r}
# Define where to save the dataset
extraWD <- tempdir()
# Get some data available to anyone
if (!file.exists(file.path(extraWD, "departement.zip"))) {
  githubURL <- "https://github.com/statnmap/blog_tips/raw/master/2018-07-14-introduction-to-mapping-with-sf-and-co/data/departement.zip"
  download.file(githubURL, file.path(extraWD, "departement.zip"))
  unzip(file.path(extraWD, "departement.zip"), exdir = extraWD)
}
```

- Reduce the dataset to a small region
```{r}
departements_l93 <- read_sf(dsn = extraWD, layer = "DEPARTEMENT")

# Reduce dataset
bretagne_l93 <- departements_l93 %>%
  filter(NOM_REG == "BRETAGNE")
```

- Calculate the regional buffer area using `regional_seas()`
```{r}
bretagne_regional_2km_l93 <- regional_seas(
  x = bretagne_l93,
  group = "NOM_DEPT",
  dist = units::set_units(30, km), # buffer distance
  density = units::set_units(0.5, 1/km) # density of points (the higher, the more precise the region attribution)
)
```

- Plot the data
```{r}
ggplot() + 
  geom_sf(data = bretagne_regional_2km_l93,
          aes(colour = NOM_DEPT, fill = NOM_DEPT),
          alpha = 0.25) +
  geom_sf(data = bretagne_l93,
          aes(fill = NOM_DEPT),
          colour = "grey20",
          alpha = 0.5) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  theme_bw()
```

